---
title: "Homework 5"
subtitle: "Research in Health Economics, Spring 2025"
author: "Lisbeth Vargas"
format:
    pdf: 
        output-file: "vargas-hwk5-1"
        output-ext: "pdf"
        header-includes: 
            - \usepackage{float}
            - \floatplacement{table}{H}
knitr:
    opts_chunk:
        warning: false

---
```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra, knitr, broom)    

library(readr)     

final.data <- read_tsv('/Users/lisbethvargas/Desktop/GitHub/Homework-5/5-1/data/output/acs_medicaid.txt')
load("/Users/lisbethvargas/Desktop/GitHub/Homework-5/5-1/analysis/hwk5_workspace.Rdata")
```

The following is my submission for Homework 5. Note that the setup and analysis for these responses are in a seperate `R` script. The GitHub repository for this work is available [here](https://github.com/lisvargasg/Homework-5.git).

\newpage
# Summarize the Data

\noindent Question 1. Plot the share of the adult population with direct purchase health insurance over time.
```{r}
#| echo=FALSE, fig.show='hold'
q1
```
\noindent Question 2. Discuss the reduction in direct purchase health insurance in later years. Can you list a couple of policies that might have affected the success of the direct purchase insurance market?

\newpage
\noindent Question 3. Plot the share of the adult population with Medicaid over time.
```{r}
#| echo=FALSE, fig.show='hold'
q3
```

\newpage
\noindent Question 4. Plot the share of uninsured over time, separately by states that expanded Medicaid in 2014 versus those that did not. Drop all states that expanded after 2014.
```{r}
#| echo=FALSE, fig.show='hold'
q4
```

\newpage
# Estimate ATEs
\noindent Question 5. Calculate the average percent of uninsured individuals in 2012 and 2015, separately for expansion and non-expansion states. Present your results in a basic 2x2 DD table.
```{r}
#| echo=FALSE, fig.show='hold'
```

\newpage
\noindent Question 6. Estimate the effect of Medicaid expansion on the uninsurance rate using a standard DD regression estimator, again focusing only on states that expanded in 2014 versus those that never expanded.
```{r}
#| echo=FALSE, fig.show='hold'
```

\newpage
\noindent Question 7. Include state and year fixed effects in your estimates. Try using the lfe or fixest package to estimate this instead of directly including the fixed effects.
```{r}
#| echo=FALSE, fig.show='hold'
```

\newpage
\noindent Question 8. Repeat the analysis in question 7 but include all states (even those that expanded after 2014). Are your results different? If so, why?
```{r}
#| echo=FALSE, fig.show='hold'
```

WORDS WORDS WORDS 